Global Life Expectancy: Unraveling Health and Economic Determinants (2015-2020)

### 2024-12-03 paper/paper.qmd
i

@fig-cleaned shows the average of life expectancy under male, female, 6 different regions and 4 income groups.

@@

```{r}
#| label: fig-cleaned
#| fig-cap: ""
#| echo: false
#| eval: true
#| warning: false
#| message: false

```
More information on the data cleaning process can be found in @sec-appx.

@@

The outcome variable is **`Life Expendency`**. This is the primary dependent variable that the model is designed to predict. It represents the average number of years a person is expected to live, under the condition that current mortality conditions persist. The model seeks to identify the variables affecting this average.

@@

# Create 10-year bins for Life Expectancy
lerange <- lelecleaned_data %>%
  mutate(LifeExpectancyRange = cut(`Life Expectancy`, 
                                   breaks = seq(floor(min(`Life Expectancy`, na.rm = TRUE)), 
                                   ceiling(max(`Life Expectancy`, na.rm = TRUE)), by = 10), 
                                   include.lowest = TRUE))

# Calculate percentages for each bin
life_expectancy_distribution <- lerange %>%
  count(LifeExpectancyRange) %>%
  mutate(Percentage = 100 * n / sum(n))

# Plot the histogram with percentages
ggplot(life_expectancy_distribution, aes(x = LifeExpectancyRange, y = Percentage, fill = LifeExpectancyRange)) +
  geom_bar(stat = "identity", color = "white", alpha = 0.8) +
  labs(
    title = "Distribution of Life Expectancy by Age Range",
    x = "Life Expectancy (10-Year Ranges)",
    y = "Percentage"
  ) +
  theme_minimal() +
  scale_fill_viridis_d(option = "C") +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

@@

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

Our results are summarized in @tbl-modelresults.

1.  check different continent has its different life expectancy.
2.  compare the develop country with
3.  check the shcooling year of develop countries to developing country, showing a relationship between shcooling with the life expenctancy.
4.

@@

@@

## First discussion point {#sec-first-point}

@@

@@

**Survey: Understanding Life Expectancy and Influencing Factor**

Thank you for participating in this survey. This survey aims to gather insights into the factors influencing life expectancy, including health expenditure, education, ethnicity, government spending on healthcare, and income levels. Your responses will help us understand individual perspectives and experiences towards national policy. Participation is voluntary, and your answers will remain anonymous.

@@

**Section 1: Personal Health Expenditure**

**1.What percentage of your monthly income do you spend on healthcare (e.g., insurance, medications, doctor visits)?**

-   Less than 5%
-   5%–10%
-   10%–20%
-   More than 20%

**2.Do you or your household have health insurance coverage?**

-   Yes
-   No

**3. How frequently do you visit a healthcare professional(eg.doctors, nurse) in a year?**

-   0-1 times
-   2-5 times
-   6-10 times
-   More than 10 times

**Section 2: Education Background**

**4.What is the highest level of education you have completed?**

-   No formal education
-   Primary school
-   Secondary school
-   College or university degree
-   Postgraduate degree

**5.How many years of formal schooling have you completed?**

Please write the number: \_\_\_\_\_\_\_

**Section 3: Ethnicity and Geographic Factors**

**6.Which of the following best describes your ethnicity?**

-   African
-   Asian
-   European
-   Hispanic or Latino
-   Middle Eastern
-   Indigenous
-   Other (please specify): \_\_\_\_\_\_\_

**7.What's the distance to the nearest healthcare facility from your residence?**

-   Less than 1 km
-   1-5 km
-   More than 5 km

**Section 4: Government Expenditure on Healthcare**

**8.Are health services in your country subsidized by the govenment?**

-   Yes, fully subsidized
-   Yes, partially subsidized
-   No

**9. What is the approximate cost of your most recent healthcare visit (in local currency)?**

Please write the amount \_\_\_\_\_\_\_

**10.Do public healthcare facilities in your area provide all the services you need?**

-   Yes
-   No
-   Not applicable

**Section 5: Income and Development Factor**

**11.What is your approximate monthly household income after taxes?(in local currency)**

-   Less than 1,000
-   1,000 - 2,999
-   3,000 - 4,999
-   5,000 - 7,999
-   8,000 - 10,999
-   More than 11,000
-   Prefer not to say

**12.How would you describe your current financial situation?**

-   Struggling to make ends meet
-   Just getting by
-   Comfortable, but not wealthy
-   Financially secure
-   Wealthy
-   Prefer not to say

**13.How often do you have difficulty affording basic healthcare (medical visits, medications, etc.)?**

-   Never
-   Rarely
-   Sometimes
-   Often
-   Always
-   Prefer not to say

>>>>

shows the average of life expectancy under male, female, 6 different regions and 4 income groups.

@@

More information on the data cleaning process can be found in @sec-appx.

@@

The outcome variable is **Life Expectancy**. This is the primary dependent variable that the model is designed to predict. It represents the average number of years a person is expected to live, under the condition that current mortality conditions persist. The model seeks to identify the variables affecting this average.

@@

# Normalize Life Expectancy to percentages
life_expectancy_percentage <- lelecleaned_data %>%
  mutate(
    LifeExpectancyPercentage = 100 * (`Life Expectancy` - min(`Life Expectancy`, na.rm = TRUE)) / 
                                  (max(`Life Expectancy`, na.rm = TRUE) - min(`Life Expectancy`, na.rm = TRUE))
  )

# Plot the distribution
ggplot(life_expectancy_percentage, aes(x = LifeExpectancyPercentage)) +
  geom_histogram(
    bins = 20, 
    fill = "skyblue", 
    color = "white", 
    alpha = 0.8
  ) +
  geom_density(
    aes(y = ..scaled.. * max(..count..)),  # Align density with histogram
    color = "blue", 
    size = 1
  ) +
  labs(
    title = "Distribution of Life Expectancy as Percentages",
    x = "Life Expectancy (%)",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

The GAM was chosen for this analysis due to its ability to capture non-linear relationships between predictors and the outcome variable, which is essential when dealing with categorical predictors like income group and the

@@

## Overview of model results
Our results are summarized in @tbl-modelresults. We are primarily interest in the

@@

## Multiple linear regression results

@@

## Insights into different gender

@@

```{r}
#| label: tbl-cleaned_leab
#| tbl-cap: "Life expectancy over countries and years"
#| echo: false
#| eval: true
#| warning: false
#| message: false

#Show the whole graph of cleaned table leab
kable(leab, 
      col.names = names(leab), 
      align = "l", 
      caption = "Life Expectancy over countries and years")

@@

**Survey: Life Expectancy and Lifestyle Survey**

Thank you for participating in this survey. This survey aims to gather insights into the factors influencing life expectancy, including gender, region and income group. Your responses will help us understand individual perspectives and experience towards national policy. Participation is voluntary, and your answers will remain anonymous.

@@

**Section 1: Demographics**

**1. Gender**

-   Male
-   Female
-   Non-binary
-   Prefer not to say

**2. Age (in years)**

Please write the number: \_\_\_\_\_\_

**3. Region of Residence**

-   Africa
-   Asia
-   Europe
-   North America
-   South America
-   Oceania

**4. Income Level**

Annual Income (Optional): \_\_\_\_\_\_ USD

Income Group:
-   Low Income (Less than $1,000/year)
-   Lower-Middle Income (\$1,001 - \$10,000/year)
-   Upper-Middle Income (\$10,001 - \$50,000/year)
-   High Income (More than \$50,000/year)

**Section 2: Health and Lifestyle**

**5. Current Life Expectancy Perception**

How many years do you expect to live?

Please write the number: \_\_\_\_\_\_

**6. Healthcare Access**

How often do you visit healthcare professionals
-   Regularly (e.g., annual checkups)
-   Occasionally (e.g., only when unwell)
-   Rarely or Never

**7. Dietary Habits**

How would you describe your diet?

-   Balanced and healthy
-   Somewhat balanced
-   Unhealthy

**8. Physical Activity**

How many hours of physical activity do you engage in per week?
-   Less than 1 hour
-   1-3 hours
-   More than 3 hours

**9. Smoking Habits**

Do you smoke?
-   Yes
-   No

**10. Alcohol Consumption**

How often do you consume alcoholic beverages?
-   Never
-   Occasionally (e.g., social drinking)
-   Frequently

**Section 3: Environmental and Social Factors**

**11.Living Environment**

How would you describe your living area?
-   Urban
-   Suburban
-   Rural
